---
title: "Lecture 4 in-class creating tidy datasets"
output:
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

# Quick review

In-class from last time:

# Problem 1 -- Data import
## Part a:
Import data from a girder strength experiment described in Wu and Hamada's Experiment Design book.  This data has shear strength data obtained by 10 different methods on 9 different girders.
```{r}
#data_file <- "https://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/fullgirder.dat"
#raw_data<-read.delim(data_file,header=T,sep = " ",stringsAsFactors = F)
#saveRDS(raw_data,file = "raw_data.RDS")
raw_data_local<-readRDS(file = "raw_data.RDS")
save.image(file = "girder.Rdata")
```
## Part b:
What if the girder names, eg "S1/1", had a meaning such as "steel girder 1 from lot 1" and you wanted to include this in your model.  Can you import the data with the name correctly parsed?  HINT: ?fread.
```{r}
library(data.table)
raw_data2<-fread(data_file,header=F,sep = "/",sep2=" ")
```

# Problem 2 -- summary functions
1. what does the imported data look like?  Missing values, odd structures?
2. What summaries, text or plots, might you make on the girder data as imported assuming 9 different girders? 

# Problem 3 -- dplyr
1. Using dplyr functions, can you summarize the data? Mean by method, mean by girder?  Min? Max?
```{r}
library(dplyr)
library(tidyr)
iris %>% 
    group_by(Species) %>%
    summarize(avg=mean(Sepal.Width)) 
```

# Problem 4 -- at home
Bitbucket config and swirl


```{r eval=F, echo=T}
install_course("Getting_and_Cleaning_Data")
```


# Don't forget the cheat sheet

<https://www.rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf>

# Example 1
Gold Medal performance for Olympic Men's Long Jump, year is coded as 1900=0.  
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/LongJumpData.dat>
```{r}
# look at first few lines
# what strategy should we use
# 
```

# Example 2
Sensory data from five operators.   
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat> 
```{r}
# look at first few lines
# what strategy should we use
# 
```


```{r Example2_Sensory_analysis, echo=F, eval=T, tidy=T, tidy.opts=list(width.cutoff=55)}
    ########################### 
    #Sensory analysis  
    #get data  
    ########################### 
    url<-"http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/Sensory.dat"
    Sensory_raw<-read.table(url, header=F, skip=1, fill=T, stringsAsFactors = F)
    saveRDS(Sensory_raw,file = "Sensory_raw.RDS")
    Sensory_raw<-readRDS(file = "Sensory_raw.RDS")
    Sensory_tidy<-Sensory_raw[-1,]
    Sensory_tidy_a<-filter(.data = Sensory_tidy,V1 %in% 1:10) %>%
                    rename(Item=V1,V1=V2,V2=V3,V3=V4,V4=V5,V5=V6)
    Sensory_tidy_b<-filter(.data = Sensory_tidy,!(V1 %in% 1:10)) %>%
                    mutate(Item=rep(as.character(1:10),each=2)) %>%
                    mutate(V1=as.numeric(V1)) %>%
                    select(c(Item,V1:V5))
    Sensory_tidy<-bind_rows(Sensory_tidy_a,Sensory_tidy_b)
    colnames(Sensory_tidy)<-c("Item",paste("Person",1:5,sep="_"))
    Sensory_tidy<-Sensory_tidy %>%  
        gather(Person,value,Person_1:Person_5) %>%  
        mutate(Person = gsub("Person_","",Person)) %>%
        arrange(Item)
    
    ########################### 
    
```

# Your turn

# Problem 1
Brain weight (g) and body weight (kg) for 62 species.    
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/BrainandBodyWeight.dat>

# Problem 2
Triplicate measurements of tomato yield for two varieties of tomatos at three planting densities.  
<http://www2.isye.gatech.edu/~jeffwu/wuhamadabook/data/tomato.dat>  
